# Tributary to Parsnip River - 125000 {-}

```{r  load-data-125000}
my_site <- 125000
```

A map of the watershed is provided in map attachment [93I.111]( https://hillcrestgeo.ca/outgoing/fishpassage/projects/parsnip/archive/2022-05-27/FishPassage_093I.111.pdf).

<br>

In 2019 a video survey was conducted by drone and can be viewed [here](https://www.youtube.com/watch?v=-foWvX1MSkg). In the summer of 2022 additional surveys were conducted with resulting images stitched into an orthomosaic and 3-dimensional model. Surveys were conducted with a remotely piloted aircraft at the crossing with resulting images stitched into an orthomosaic and 3-dimensional model `r if(gitbook_on){knitr::asis_output(" presented in Figures \\@ref(fig:ortho-caption-arctic) - \\@ref(fig:model-caption-arctic).")} else knitr::asis_output("presented [here](https://www.mapsmadeeasy.com/maps/public/0b96f06f52be467db7250b3caeeecee3) and [here](https://www.mapsmadeeasy.com/maps/public_3D/0b96f06f52be467db7250b3caeeecee3/).")`

```{r ortho-arctic, eval = identical(gitbook_on, TRUE)}
model_url <- '<iframe src="https://www.mapsmadeeasy.com/maps/public/0b96f06f52be467db7250b3caeeecee3" scrolling="no" title="Maps Made Easy" width="80%" height="400" frameBorder ="0"></iframe>'

knitr::asis_output(model_url)
```


```{r ortho-caption-arctic, fig.cap= my_caption, out.width = "0.01%", eval = identical(gitbook_on, TRUE)}
my_photo = 'fig/pixel.png'

my_caption = paste0('Orthomosaic of site of planned bridge on tributary to Parsnip River. To zoom press "shift" and scroll.')

knitr::include_graphics(my_photo, dpi = NA)

```

<br>

```{r model-arctic, eval = identical(gitbook_on, TRUE)}
model_url <- '<iframe src="https://www.mapsmadeeasy.com/maps/public_3D/0b96f06f52be467db7250b3caeeecee3" scrolling="no" title="Maps Made Easy" width="80%" height="400" frameBorder ="0"></iframe>'

knitr::asis_output(model_url)
```


```{r model-caption-arctic, fig.cap= my_caption, out.width = "0.01%", eval = identical(gitbook_on, TRUE)}
my_photo = 'fig/pixel.png'

my_caption = paste0('3D model of crossing 125000 where newly installed bridge is planned on a tributary to Parsnip River. To zoom press "shift" and scroll.')

knitr::include_graphics(my_photo, dpi = NA)

```



